Investigating Determinism of Physics Engines for usage in CAV and V\&V simulations

### 2021-04-14 document/source/body.tex
TP comments complete

The experiment design is based around a series of carefully selected tests that varies one of the sources of non-determinism whilst keeping all others parameters constant. Each test is repeated $n$ times and the analysis of the results provide a confidence with a degree of statistical certainty. By varying a single parameter and controlling all others, the simulation variance associated with each source of non-determinism can be found and addressed.

@@

\subsubsection{Simulation Time} 
The simulation time will depend on the actor paths and terminating conditions. The simulation time should be sufficient to record interactions between actors but not so long that the testing take an inconvenient time to complete. In our empirical investigation a simulation time of $10-20$s was sufficient to monitor the distinct change in events such as the pre- and post-collision including the delayed effect seen by other actors, shown in Fig.~\ref{CarsPeopleCollsionsCG25} in Section~\ref{s:FinalResultsSection}. The termination conditions of the simulation can be set by, for example, actors reaching their final trajectory waypoints.

@@

\subsubsection{Scheduling}
We hypothesise that thread scheduling may be a major contributor to the non-deterministic results found in the empirical study. However, gaining fine control over the scheduling policy and thread execution order is known to not be trivial~\cite{acm-q-rr-interview}.

@@

\subsubsection{Ego Vehicle Controller} 
An ego vehicle was not used in this study, but the impact of introducing this to the simulator can be considered here. The ego vehicle is seen as another actor in the simulation but care must be taken to ensure that any control algorithms, machine learning modules and processing pipelines are deterministic. We recommended this be treated as a separate source of non-determinism and handled accordingly.

@@

\subsubsection{Sample Size}
It is recommended that the chosen sample size is determined empirically. We recommend monitoring variance while increasing the sample size in orders of magnitude until there is confidence that $\max\sigma$ will not exceed the permissible variance for the verification process. Note that we suggest monitoring the maximum value of $\sigma$, not the average, because if even a single simulation run is outside of the permissible variance this may lead to false confidence in the verification result.

@@

A general method to assess the actor path variance of a game-engine-based simulation environment is then proposed. The method can be used to find the \textit{domains of permissible variance} of a simulation environment for a given system configuration. This can give AV developers and verification engineers increased confidence in the simulation results and reduce debug time. As future work the method can be extended to criteria other than the actor path, e.g.\ actor orientation and any status indicators that may be of interest, also including actions, sequences and timings that may be useful for verification purposes.

>>>>

The experiment design is based around a series of carefully selected tests that varies one of the sources of non-determinism whilst keeping all other parameters constant. Each test is repeated $n$ times and the analysis of the results provide a confidence with a degree of statistical certainty. By varying a single parameter and controlling all others, the simulation variance associated with each source of non-determinism can be found and addressed.

@@

\subsubsection{Simulation Time} 
The simulation time will depend on the actor paths and terminating conditions. The simulation time should be sufficient to record interactions between actors but not so long that the testing takes an inconvenient time to complete. In our empirical investigation a simulation time of $10-20$s was sufficient to monitor the distinct change in events such as the pre- and post-collision including the delayed effect seen by other actors, shown in Fig.~\ref{CarsPeopleCollsionsCG25} in Section~\ref{s:FinalResultsSection}. The termination conditions of the simulation can be set by, for example, actors reaching their final trajectory waypoints.

@@

\subsubsection{Scheduling}
We hypothesise that thread scheduling may be a major contributor to the non-deterministic results found in the empirical study. However, gaining fine control over the scheduling policy and thread execution order is non-trivial~\cite{acm-q-rr-interview}.

@@

\subsubsection{Ego Vehicle Controller} 
An ego vehicle was not used in this study, but the impact of introducing this to the simulator can be considered here. The ego vehicle is seen as another actor in the simulation but care must be taken to ensure that any control algorithms, machine learning modules and processing pipelines are deterministic. We recommended that this be treated as a separate source of non-determinism and handled accordingly.

@@

\subsubsection{Sample Size}
It is recommended that the chosen sample size, i.e. the number of repeated tests, is determined empirically. We recommend monitoring variance while increasing the sample size in orders of magnitude until there is confidence that $\max\sigma$ will not exceed the permissible variance for the verification process. Note that we suggest monitoring the maximum value of $\sigma$, not the average, because if even a single simulation run is outside of the permissible variance this may lead to false confidence in the verification result.

@@

A general method to assess the actor path variance of a game-engine-based simulation environment was then proposed. The method can be used to find the \textit{domains of permissible variance} of a simulation environment for a given system configuration. This can give AV developers and verification engineers increased confidence in the simulation results and reduce debug time. As future work, the method can be extended to other simulation platforms and to criteria other than the actor path, e.g.\ actor orientation and any status indicators that may be of interest, also including actions, sequences and timings that may be useful for verification purposes.
